---
title: "From Summer Facilitation to Winter Avoidance: Seasonal Shifts in Livestock-Wild Ungulate Temporal Coexistence"
author: "Annika Zuleger & Melissa Einsele"
output:
  html_document:
    df_print: paged
---

#### Skript to run analysis to assess spatiotemporal effects of livestock (cattle and horses) on wild ungulates (roe deer and wild boar)

## Load required packages

```{r, message=FALSE , results='hide', warning=FALSE}
# For activity pattern analysis
library(activity)
library(overlap)
library(suncalc)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

# For GLMM spatiotemporal analysis
library(lme4)
library(effects)
library(DHARMa)
library(car)
library(emmeans)
library(dplyr)
library(spdep)
```

### Activity pattern and overlap analysis

## Load data

```{r}

## Load required data
setwd("H:/Spatiotemporal_analysis/Data")

all_occ <- read.csv("CT_Peneda_data_new.csv",header=T,sep=";") # Read dataset
all_occ$deploymentStart <- as.Date(all_occ$deploymentStart, format = "%d.%m.%Y")
all_occ$deploymentEnd <- as.Date(all_occ$deploymentEnd, format = "%d.%m.%Y")
all_occ$eventStartTime <- strptime(all_occ$eventStartTime, format = "%H:%M:%S")

# Calculate sampling effort (camera trap days per location)
for (i in 1:nrow(all_occ)) {
  if(year(all_occ$deploymentStart)[i] == year(all_occ$deploymentEnd)[i]){
    all_occ$sampleSizeValue[i] <- as.numeric((all_occ$deploymentEnd[i]+1) - all_occ$deploymentStart[i])
  } else {
    all_occ$sampleSizeValue[i] <- as.numeric((all_occ$deploymentEnd[i] + 365) - all_occ$deploymentStart[i])
  }
}

uni_loc <- all_occ[!duplicated(all_occ[,c('locationID','Season')]),]

uni_loc_summary <- uni_loc %>%
  group_by(Season) %>%
  summarise(
    Sampling_effort = sum(sampleSizeValue),
    .groups = "drop"
  )

## Create dataset for analysis
species <- c("Bos taurus", "Equus caballus", "Canis lupus familiaris", "Capreolus capreolus", "Sus scrofa", "Capra pyrenaica", "Canis lupus")

# Subset the dataset
all_occ <- all_occ[all_occ$scientificName %in% species, ]

# Get species specific counts
table(all_occ$scientificName)

# Calculate observation frequencies and encounter rates for cattle and horses
freq_er_summary <- all_occ %>%
  group_by(locationID, sampleSizeValue) %>%
  summarise(
    Cattle_freq = sum(scientificName == "Bos taurus", na.rm = TRUE),
    Horse_freq = sum(scientificName == "Equus caballus", na.rm = TRUE),
    Dog_freq = sum(scientificName == "Canis lupus familiaris", na.rm = TRUE),
    Deer_freq = sum(scientificName == "Capreolus capreolus", na.rm = TRUE),
    Boar_freq = sum(scientificName == "Sus scrofa", na.rm = TRUE),
    Ibex_freq = sum(scientificName == "Capra pyrenaica", na.rm = TRUE),
    Wolf_freq = sum(scientificName == "Canis lupus", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Cattle_er = Cattle_freq / sampleSizeValue,
    Horse_er = Horse_freq / sampleSizeValue,
    Dog_er = Dog_freq / sampleSizeValue,
    Deer_er = Deer_freq / sampleSizeValue,
    Boar_er = Boar_freq / sampleSizeValue,
    Ibex_er = Ibex_freq / sampleSizeValue,
    Wolf_er = Wolf_freq / sampleSizeValue
  )

print(freq_er_summary)

lm_dog_cattle <- lm(Dog_er ~ Cattle_er,freq_er_summary)
summary(lm_dog_cattle)

plot(Dog_er ~ Cattle_er,freq_er_summary)
abline(lm_dog_cattle,col="red")

all_occ <- merge(all_occ,freq_er_summary, by = c("locationID", "sampleSizeValue"), all.x = T)

total_er <- all_occ %>%
  distinct(locationID, .keep_all = TRUE) %>%
  pivot_longer(
    cols = ends_with("_er"),
    names_to = "species",
    values_to = "value"
  ) %>%
  group_by(species) %>%
  summarise(
    mean   = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min    = min(value, na.rm = TRUE),
    max    = max(value, na.rm = TRUE),
    sd     = sd(value, na.rm = TRUE),
    .groups = "drop"
  )

total_er

# Create factor variable for high and low abundance
all_occ$Cattle <- ifelse(all_occ$Cattle_er > total_er$mean[total_er$species == "Cattle_er"],"high", "low")
all_occ$Horse <- ifelse(all_occ$Horse_er > total_er$mean[total_er$species == "Horse_er"], "high", "low")
all_occ$Dog <- ifelse(all_occ$Dog_er > 0, 1, 0)

## Get average sunrise and sunset times for summer and winter
dates_summer <- seq(as.Date("2021-04-01"), as.Date("2021-09-30"), by="days") # create vector of dates for summer
dates_summer<- getSunlightTimes(dates_summer,42.01683, -8.184782, keep= c("sunrise","sunset"),tz="Europe/London") # obtain sunrise and sunset times for each date

mean.POSIXct(dates_summer$sunrise) # estimate mean sunrise time
mean.POSIXct(dates_summer$sunset)  # estimate mean sunset time

dates_winter <- seq(as.Date("2021-10-01"), as.Date("2022-03-31"), by="days")
dates_winter <- getSunlightTimes(dates_winter,42.01683, -8.184782, keep= c("sunrise","sunset"),tz="Europe/Berlin")

mean.POSIXct(dates_winter$sunrise)
mean.POSIXct(dates_winter$sunset)

# Summer data only
summer <- subset(all_occ,Season=="summer") # create a subset of observations in summer

table(summer$scientificName) # Get species counts

summer_er <- summer %>%
  distinct(locationID, .keep_all = TRUE) %>%
  pivot_longer(
    cols = ends_with("_er"),
    names_to = "species",
    values_to = "value"
  ) %>%
  group_by(species) %>%
  summarise(
    mean   = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min    = min(value, na.rm = TRUE),
    max    = max(value, na.rm = TRUE),
    sd     = sd(value, na.rm = TRUE),
    .groups = "drop"
  )

summer_er

summer$Cattle <- ifelse(summer$Cattle_er > summer_er$mean[summer_er$species == "Cattle_er"],"high", "low")
summer$Horse <- ifelse(summer$Horse_er > summer_er$mean[summer_er$species == "Horse_er"], "high", "low")
summer$Dog <- ifelse(summer$Dog_er > 0, 1, 0)

# Winter data only 
winter <- subset(all_occ,Season=="winter") # create a subset of observations in winter

table(winter$scientificName) # Get species counts

winter_er <- winter %>%
  distinct(locationID, .keep_all = TRUE) %>%
  pivot_longer(
    cols = ends_with("_er"),
    names_to = "species",
    values_to = "value"
  ) %>%
  group_by(species) %>%
  summarise(
    mean   = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min    = min(value, na.rm = TRUE),
    max    = max(value, na.rm = TRUE),
    sd     = sd(value, na.rm = TRUE),
    .groups = "drop"
  )

winter_er

winter$Cattle <- ifelse(winter$Cattle_er > winter_er$mean[winter_er$species == "Cattle_er"],"high", "low")
winter$Horse <- ifelse(winter$Horse_er > winter_er$mean[winter_er$species == "Horse_er"], "high", "low")
winter$Dog <- ifelse(winter$Dog_er > 0, 1, 0)

```

## Define function to estimate diurnal and nocturnal activity levels (provided by M. Rowcliffe)

```{r}
# Function takes a fitted 24-hour activity model (fit) and a 2-element vector (range) giving the radian times between which to calculate the percentage of time spent active.

partial.act <- function(fit, range){
  pd <- head(fit@pdf[,2], -1) #extract prob densities and matching times (minus final overlapping value)
  tt <- head(fit@pdf[,1], -1)
  if(range[1]>range[2]){ #if range overlaps midnight...
    tt[tt<range[2]] <- tt[tt<range[2]]+2*pi #increment lower times and second range value to make range increasing
    range[2] <- range[2]+2*pi
  }
  i <- tt>range[1] & tt<range[2] #indicator, are times within the range?
  sum(pd[i])*2*pi / (512 * (max(pd)*diff(range))) #return activity level; note fit has 512 pdf values, so the approx. area under the curve is sum(pd)*2*pi/512
}
```

### Seasonal analysis per species

## Cattle

```{r}
# Create subset of all deer observations in summer
cattle_summer <- subset(summer, scientificName == "Bos taurus")

# Fit activity function
cattle_all_summer <- solartime(cattle_summer$eventStartTime, 42.01683, -8.184782, 0) # transform times to solar times using average anchoring approach (Vazquez et al. 2019)
cattleallsummer <- fitact(cattle_all_summer$solar,sample="data",reps=100) # Fit kernel density function to estimate activity level
cattle_summer_day<-partial.act(cattleallsummer, c(6+32/60, 20+36/60)*pi/12) # estimate diurnal activity level 
cattle_summer_night<-partial.act(cattleallsummer, c(20+36/60,6+32/60)*pi/12) # estimate nocturnal activity level

# Create subset of all cattle observations in winter
cattle_winter <- subset(winter, scientificName == "Bos taurus")

# Fit activity function
cattle_all_winter <- solartime(cattle_winter$eventStartTime, 42.01683, -8.184782, 0)
cattleallwinter <- fitact(cattle_all_winter$solar,sample="data",reps=100)
cattle_winter_day<-partial.act(cattleallwinter, c(8+22/60, 18+44/60)*pi/12)
cattle_winter_night<-partial.act(cattleallwinter, c(18+44/60,8+22/60)*pi/12)

# Estimate overlap and difference in activity patterns
cattle_winter_summer_Ckern <- compareCkern(cattleallsummer, cattleallwinter, reps = 100)
cattle_winter_summer_Act <- compareAct(list(cattleallsummer, cattleallwinter))

overlapPlot(cattle_all_summer$solar, cattle_all_winter$solar, type = "Dhat4",linet = c(1,1), linewidth = c(2, 2), linec = c("orangered2", "turquoise4"),main="A) Cattle", ylim=c(0,0.12), extend="azure2", cex.lab=1.5, cex.axis=1.3,cex.main=2)
abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2,col="orangered2")
abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2,col="turquoise4")
legend("topleft", 
       c(paste0("Summer (n = ", nrow(cattle_summer),")"),
         paste0("Winter (n = ", nrow(cattle_winter),")")),
       lty=1, lwd=2,col=c("orangered2", "turquoise4"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", cattle_winter_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", cattle_winter_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)
```

## Dog

```{r}
# Create subset of all dog observations in summer
dog_summer <- subset(summer, scientificName == "Canis lupus familiaris")

dog_all_summer <- solartime(dog_summer$eventStartTime, 42.01683, -8.184782, 0) # transform times to solar times using average anchoring approach (Vazquez et al. 2019)
dogallsummer <- fitact(dog_all_summer$solar,sample="data",reps=100) # Fit kernel density function to estimate activity level
dog_summer_day<-partial.act(dogallsummer, c(6+32/60, 20+36/60)*pi/12) # estimate diurnal activity level 
dog_summer_night<-partial.act(dogallsummer, c(20+36/60,6+32/60)*pi/12) # estimate nocturnal activity level 

# Create subset of all dog observations in winter
dog_winter <- subset(winter, scientificName == "Canis lupus familiaris")

dog_all_winter <- solartime(dog_winter$eventStartTime, 42.01683, -8.184782, 0)
dogallwinter <- fitact(dog_all_winter$solar,sample="data",reps=100)
dog_winter_day<-partial.act(dogallwinter, c(8+22/60, 18+44/60)*pi/12)
dog_winter_night<-partial.act(dogallwinter, c(18+44/60,8+22/60)*pi/12)

# Estimate overlap and difference in activity patterns
dog_winter_summer_Ckern <- compareCkern(dogallsummer, dogallwinter, reps = 100)
dog_winter_summer_Act <- compareAct(list(dogallsummer, dogallwinter)) 


overlapPlot(dog_all_summer$solar, dog_all_winter$solar, type = "Dhat4",linet = c(1,1), linewidth = c(2, 2), linec = c("orangered2", "turquoise4"),main="Dog", ylim=c(0,0.2), extend="azure2", cex.lab=1.5, cex.axis=1.3,cex.main=2)
abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2,col="orangered2")
abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2,col="turquoise4")
legend("topleft", 
       c(paste0("Summer (n = ", nrow(dog_summer),")"),
         paste0("Winter (n = ", nrow(dog_winter),")")),
       lty=1, lwd=2,col=c("orangered2", "turquoise4"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", dog_winter_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", dog_winter_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

```

## Horse 

```{r}
# Create subset of all deer observations in summer
horse_summer <- subset(summer, scientificName == "Equus caballus")

horse_all_summer <- solartime(horse_summer$eventStartTime, 42.01683, -8.184782, 0) # transform times to solar times using average anchoring approach (Vazquez et al. 2019)
horseallsummer <- fitact(horse_all_summer$solar,sample="data",reps=100) # Fit kernel density function to estimate activity level
horse_summer_day<-partial.act(horseallsummer, c(6+32/60, 20+36/60)*pi/12) # estimate diurnal activity level
horse_summer_night<-partial.act(horseallsummer, c(20+36/60,6+32/60)*pi/12) # estimate nocturnal activity level 

# Create subset of all horse observations in winter
horse_winter <- subset(winter, scientificName == "Equus caballus")

horse_all_winter <- solartime(horse_winter$eventStartTime, 42.01683, -8.184782, 0)
horseallwinter <- fitact(horse_all_winter$solar,sample="data",reps=100)
horse_winter_day<-partial.act(horseallwinter, c(8+22/60, 18+44/60)*pi/12)
horse_winter_night<-partial.act(horseallwinter, c(18+44/60,8+22/60)*pi/12)

# Estimate overlap and difference in activity patterns
horse_winter_summer_Ckern <- compareCkern(horseallsummer, horseallwinter, reps = 100)
horse_winter_summer_Act <- compareAct(list(horseallsummer, horseallwinter)) 

overlapPlot(horse_all_summer$solar, horse_all_winter$solar, type = "Dhat4",linet = c(1,1), linewidth = c(2, 2), linec = c("orangered2", "turquoise4"),main="B) Horse", ylim=c(0,0.12), extend="azure2", cex.lab=1.5, cex.axis=1.3,cex.main=2)
abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2,col="orangered2")
abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2,col="turquoise4")
legend("topleft", 
       c(paste0("Summer (n = ", nrow(horse_summer),")"),
         paste0("Winter (n = ", nrow(horse_winter),")")),
       lty=1, lwd=2,col=c("orangered2", "turquoise4"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", horse_winter_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", horse_winter_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

```

## Roe deer 

```{r}
# Create subset of all deer observations in summer
deer_summer <- subset(summer, scientificName == "Capreolus capreolus")

deer_all_summer <- solartime(deer_summer$eventStartTime, 42.01683, -8.184782, 0) # transform times to solar times using average anchoring approach (Vazquez et al. 2019)
deerallsummer <- fitact(deer_all_summer$solar,sample="data",reps=100) # Fit kernel density function to estimate activity level
deer_summer_day<-partial.act(deerallsummer, c(6+32/60, 20+36/60)*pi/12) # estimate diurnal activity level 
deer_summer_night<-partial.act(deerallsummer, c(20+36/60,6+32/60)*pi/12) # estimate nocturnal activity level 

# Create subset of all deer observations in winter
deer_winter <- subset(winter, scientificName == "Capreolus capreolus")

deer_all_winter <- solartime(deer_winter$eventStartTime, 42.01683, -8.184782, 0)
deerallwinter <- fitact(deer_all_winter$solar,sample="data",reps=100)
deer_winter_day<-partial.act(deerallwinter, c(8+22/60, 18+44/60)*pi/12)
deer_winter_night<-partial.act(deerallwinter, c(18+44/60,8+22/60)*pi/12)

# Estimate overlap and difference in activity patterns
deer_winter_summer_Ckern <- compareCkern(deerallsummer, deerallwinter, reps = 100)
deer_winter_summer_Act <- compareAct(list(deerallsummer, deerallwinter)) 

overlapPlot(deer_all_summer$solar, deer_all_winter$solar, type = "Dhat4",linet = c(1,1), linewidth = c(2, 2), linec = c("orangered2", "turquoise4"),main="C) Roe deer", ylim=c(0,0.12), extend="azure2", cex.lab=1.5, cex.axis=1.3,cex.main=2)
abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2,col="orangered2")
abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2,col="turquoise4")
legend("topleft", 
       c(paste0("Summer (n = ", nrow(deer_summer),")"),
         paste0("Winter (n = ", nrow(deer_winter),")")),
       lty=1, lwd=2,col=c("orangered2", "turquoise4"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", deer_winter_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_winter_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)
```

## Wild boar 

```{r}
# Create subset of all deer observations in summer
boar_summer <- subset(summer, scientificName == "Sus scrofa")

boar_all_summer <- solartime(boar_summer$eventStartTime, 42.01683, -8.184782, 0) # transform times to solar times using average anchoring approach (Vazquez et al. 2019)
boarallsummer <- fitact(boar_all_summer$solar,sample="data",reps=100) # Fit kernel density function to estimate activity level
boar_summer_day<-partial.act(boarallsummer, c(6+32/60, 20+36/60)*pi/12) # estimate diurnal activity level
boar_summer_night<-partial.act(boarallsummer, c(20+36/60,6+32/60)*pi/12) # estimate nocturnal activity level

# Create subset of all boar observations in winter
boar_winter <- subset(winter, scientificName == "Sus scrofa")

# Estimate overlap and difference in activity patterns
boar_all_winter <- solartime(boar_winter$eventStartTime, 42.01683, -8.184782, 0)
boarallwinter <- fitact(boar_all_winter$solar,sample="data",reps=100)
boar_winter_day<-partial.act(boarallwinter, c(8+22/60, 18+44/60)*pi/12)
boar_winter_night<-partial.act(boarallwinter, c(18+44/60,8+22/60)*pi/12)

boar_winter_summer_Ckern <- compareCkern(boarallsummer, boarallwinter, reps = 100)
boar_winter_summer_Act <- compareAct(list(boarallsummer, boarallwinter)) 

overlapPlot(boar_all_summer$solar, boar_all_winter$solar, type = "Dhat4",linet = c(1,1), linewidth = c(2, 2), linec = c("orangered2", "turquoise4"),main="D) Wild boar", ylim=c(0,0.12), extend="azure2", cex.lab=1.5, cex.axis=1.3,cex.main=2)
abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2,col="orangered2")
abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2,col="turquoise4")
legend("topleft", 
       c(paste0("Summer (n = ", nrow(boar_summer),")"),
         paste0("Winter (n = ", nrow(boar_winter),")")),
       lty=1, lwd=2,col=c("orangered2", "turquoise4"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", boar_winter_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_winter_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)
```

### Effects of cattle

## Horse

```{r}
## Cattle - Summer 

# High cattle abundance
horse_cattle_summer <- subset(horse_summer, Cattle == "high") # Create data subset
horse_cattle_all_summer <- solartime(horse_cattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsecattlesummer <- fitact(horse_cattle_all_summer$solar,sample="data",reps=100) # Fit activity function
horse_cattle_summer_day<-partial.act(horsecattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_cattle_summer_night<-partial.act(horsecattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
horse_nocattle_summer  <- subset(horse_summer, Cattle == "low") # Create data subset
horse_nocattle_all_summer  <- solartime(horse_nocattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsenocattlesummer <- fitact(horse_nocattle_all_summer$solar,sample="data",reps=100) # Fit activity function
horse_nocattle_summer_day<-partial.act(horsenocattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_nocattle_summer_night<-partial.act(horsenocattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
horse_cattle_summer_Ckern <- compareCkern(horsecattlesummer, horsenocattlesummer, reps = 100) # Overlap activity patterns
horse_cattle_summer_Act<- compareAct(list(horsecattlesummer, horsenocattlesummer)) # Difference in activity levels

# Plot
overlap::densityPlot(cattle_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="A) Horse - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_summer$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(horse_cattle_all_summer $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(horse_nocattle_all_summer $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", 
       c(paste0("Cattle High (n = ", nrow(horse_cattle_summer),")"),
         paste0("Cattle Low (n = ", nrow(horse_nocattle_summer),")"),
         paste0("Cattle (n = ", nrow(cattle_summer),")"),
         paste0("Dog (n = ", nrow(dog_summer),")")),
       lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", horse_cattle_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", horse_cattle_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
horse_cattle_summer_direct_Ckern <- compareCkern(horsecattlesummer, cattleallsummer, reps = 100)
horse_nocattle_summer_direct_Ckern <- compareCkern(horsenocattlesummer, cattleallsummer, reps = 100)

## Dog - Summer 

# Dog presence
horse_dog_summer <- subset(horse_summer, Dog == 1) # Create data subset
horse_dog_all_summer <- solartime(horse_dog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsedogsummer <- fitact(horse_dog_all_summer$solar,sample="data",reps=100) # Fit activity function
horse_dog_summer_day<-partial.act(horsedogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_dog_summer_night<-partial.act(horsedogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
horse_nodog_summer  <- subset(horse_summer, Dog == 0) # Create data subset
horse_nodog_all_summer  <- solartime(horse_nodog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsenodogsummer <- fitact(horse_nodog_all_summer$solar,sample="data",reps=100) # Fit activity function
horse_nodog_summer_day<-partial.act(horsenodogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_nodog_summer_night<-partial.act(horsenodogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
horse_dog_summer_Ckern <- compareCkern(horsedogsummer, horsenodogsummer, reps = 100) # Overlap activity patterns
horse_dog_summer_Act<- compareAct(list(horsedogsummer, horsenodogsummer)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="A) Horse - Dog - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(horse_dog_all_summer $solar,  add=T, n.grid=100, lwd=2, col="coral2")
overlap::densityPlot(horse_nodog_all_summer $solar, add=T, n.grid=100,col="cyan2",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("coral2","cyan2", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", horse_dog_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", horse_dog_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
horse_dog_summer_direct_Ckern <- compareCkern(horsedogsummer, dogallsummer, reps = 100)
horse_nodog_summer_direct_Ckern <- compareCkern(horsenodogsummer, dogallsummer, reps = 100)

## Cattle - Winter

# High cattle abundance
horse_cattle_winter <- subset(horse_winter, Cattle == "high") # Create data subset
horse_cattle_all_winter <- solartime(horse_cattle_winter$eventStartTime, 42.01683, -8.184782, 0)# Transform to solar time
horsecattlewinter <- fitact(horse_cattle_all_winter$solar,sample="data",reps=100) # Fit activity function
horse_cattle_winter_day<-partial.act(horsecattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_cattle_winter_night<-partial.act(horsecattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
horse_nocattle_winter  <- subset(horse_winter, Cattle == "low") # Create data subset
horse_nocattle_all_winter  <- solartime(horse_nocattle_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsenocattlewinter <- fitact(horse_nocattle_all_winter$solar,sample="data",reps=100) # Fit activity function
horse_nocattle_winter_day<-partial.act(horsenocattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_nocattle_winter_night<-partial.act(horsenocattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
horse_cattle_winter_Ckern <- compareCkern(horsecattlewinter, horsenocattlewinter, reps = 100) # Overlap activity patterns
horse_cattle_winter_Act <- compareAct(list(horsecattlewinter, horsenocattlewinter)) # Difference in activity levels

# Plot
overlap::densityPlot(cattle_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="B) Horse - Winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_winter$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(horse_cattle_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(horse_nocattle_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2)
legend("topleft", c("High cattle abundances","Low cattle abundances", "Cattle", "Shepherd dog"), lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", horse_cattle_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", horse_cattle_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
horse_cattle_winter_direct_Ckern <- compareCkern(horsecattlewinter, cattleallwinter, reps = 100)
horse_nocattle_winter_direct_Ckern <- compareCkern(horsenocattlewinter, cattleallwinter, reps = 100)

## Dog - Winter

# Dog presence
horse_dog_winter <- subset(horse_winter, Dog == 1) # Create data subset
horse_dog_all_winter <- solartime(horse_dog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsedogwinter <- fitact(horse_dog_all_winter$solar,sample="data",reps=100) # Fit activity function
horse_dog_winter_day<-partial.act(horsedogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_dog_winter_night<-partial.act(horsedogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
horse_nodog_winter  <- subset(horse_winter, Dog == 0) # Create data subset
horse_nodog_all_winter  <- solartime(horse_nodog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
horsenodogwinter <- fitact(horse_nodog_all_winter$solar,sample="data",reps=100) # Fit activity function
horse_nodog_winter_day<-partial.act(horsenodogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
horse_nodog_winter_night<-partial.act(horsenodogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
horse_dog_winter_Ckern <- compareCkern(horsedogwinter, horsenodogwinter, reps = 100) # Overlap activity patterns
horse_dog_winter_Act<- compareAct(list(horsedogwinter, horsenodogwinter)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="A) Horse - Dog - winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(horse_dog_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(horse_nodog_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("purple3","mediumseagreen", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", horse_dog_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", horse_dog_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
horse_dog_winter_direct_Ckern <- compareCkern(horsedogwinter, dogallwinter, reps = 100)
horse_nodog_winter_direct_Ckern <- compareCkern(horsenodogwinter, dogallwinter, reps = 100)

```

## Roe deer

```{r}
## Summer 

# High cattle abundance
deer_cattle_summer <- subset(deer_summer, Cattle == "high") # Create data subset
deer_cattle_all_summer <- solartime(deer_cattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deercattlesummer <- fitact(deer_cattle_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_cattle_summer_day<-partial.act(deercattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_cattle_summer_night<-partial.act(deercattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
deer_nocattle_summer  <- subset(deer_summer, Cattle == "low") # Create data subset
deer_nocattle_all_summer  <- solartime(deer_nocattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernocattlesummer <- fitact(deer_nocattle_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_nocattle_summer_day<-partial.act(deernocattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nocattle_summer_night<-partial.act(deernocattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_cattle_summer_Ckern <- compareCkern(deercattlesummer, deernocattlesummer, reps = 100) # Overlap activity patterns
deer_cattle_summer_Act<- compareAct(list(deercattlesummer, deernocattlesummer)) # Difference in activity levels 

# Plot
overlap::densityPlot(cattle_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="A) Roe deer - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_summer$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(deer_cattle_all_summer $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(deer_nocattle_all_summer $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("High cattle abundances","Low cattle abundances", "Cattle", "Shepherd dog"), lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", deer_cattle_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_cattle_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
deer_cattle_summer_direct_Ckern <- compareCkern(deercattlesummer, cattleallsummer, reps = 100)
deer_nocattle_summer_direct_Ckern <- compareCkern(deernocattlesummer, cattleallsummer, reps = 100)


# Dog presence
deer_dog_summer <- subset(deer_summer, Dog == 1) # Create data subset
deer_dog_all_summer <- solartime(deer_dog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deerdogsummer <- fitact(deer_dog_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_dog_summer_day<-partial.act(deerdogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_dog_summer_night<-partial.act(deerdogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
deer_nodog_summer  <- subset(deer_summer, Dog == 0) # Create data subset
deer_nodog_all_summer  <- solartime(deer_nodog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernodogsummer <- fitact(deer_nodog_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_nodog_summer_day<-partial.act(deernodogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nodog_summer_night<-partial.act(deernodogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_dog_summer_Ckern <- compareCkern(deerdogsummer, deernodogsummer, reps = 100) # Overlap activity patterns
deer_dog_summer_Act<- compareAct(list(deerdogsummer, deernodogsummer)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="A) Roe deer - Dog - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(deer_dog_all_summer $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(deer_nodog_all_summer $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("purple3","mediumseagreen", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", deer_dog_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_dog_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
deer_dog_summer_direct_Ckern <- compareCkern(deerdogsummer, dogallsummer, reps = 100)
deer_nodog_summer_direct_Ckern <- compareCkern(deernodogsummer, dogallsummer, reps = 100)

## Winter 

# High cattle abundance
deer_cattle_winter <- subset(deer_winter, Cattle == "high") # Create data subset
deer_cattle_all_winter <- solartime(deer_cattle_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deercattlewinter <- fitact(deer_cattle_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_cattle_winter_day<-partial.act(deercattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_cattle_winter_night<-partial.act(deercattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
deer_nocattle_winter  <- subset(deer_winter, Cattle == "low") # Create data subset
deer_nocattle_all_winter  <- solartime(deer_nocattle_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernocattlewinter <- fitact(deer_nocattle_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_nocattle_winter_day<-partial.act(deernocattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nocattle_winter_night<-partial.act(deernocattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_cattle_winter_Ckern <- compareCkern(deercattlewinter, deernocattlewinter, reps = 100) # Overlap activity patterns
deer_cattle_winter_Act <- compareAct(list(deercattlewinter, deernocattlewinter)) # Difference in activity levels 

# Plot
overlap::densityPlot(cattle_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="B) Roe deer - Winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_winter$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(deer_cattle_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(deer_nocattle_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2)
legend("topleft", c("High cattle abundances","Low cattle abundances", "Cattle", "Shepherd dog"), lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", deer_cattle_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_cattle_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
deer_cattle_winter_direct_Ckern <- compareCkern(deercattlewinter, cattleallwinter, reps = 100)
deer_nocattle_winter_direct_Ckern <- compareCkern(deernocattlewinter, cattleallwinter, reps = 100)

# Dog presence
deer_dog_winter <- subset(deer_winter, Dog == 1) # Create data subset
deer_dog_all_winter <- solartime(deer_dog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deerdogwinter <- fitact(deer_dog_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_dog_winter_day<-partial.act(deerdogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_dog_winter_night<-partial.act(deerdogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
deer_nodog_winter  <- subset(deer_winter, Dog == 0) # Create data subset
deer_nodog_all_winter  <- solartime(deer_nodog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernodogwinter <- fitact(deer_nodog_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_nodog_winter_day<-partial.act(deernodogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nodog_winter_night<-partial.act(deernodogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_dog_winter_Ckern <- compareCkern(deerdogwinter, deernodogwinter, reps = 100) # Overlap activity patterns
deer_dog_winter_Act<- compareAct(list(deerdogwinter, deernodogwinter)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="A) Roe deer - Dog - winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(deer_dog_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(deer_nodog_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("purple3","mediumseagreen", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", deer_dog_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_dog_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
deer_dog_winter_direct_Ckern <- compareCkern(deerdogwinter, dogallwinter, reps = 100)
deer_nodog_winter_direct_Ckern <- compareCkern(deernodogwinter, dogallwinter, reps = 100)

```

## Wild boar

```{r}
## Summer

# High cattle abundance
boar_cattle_summer <- subset(boar_summer, Cattle == "high") # Create data subset
boar_cattle_all_summer <- solartime(boar_cattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarcattlesummer <- fitact(boar_cattle_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_cattle_summer_day<-partial.act(boarcattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_cattle_summer_night<-partial.act(boarcattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
boar_nocattle_summer  <- subset(boar_summer, Cattle == "low") # Create data subset
boar_nocattle_all_summer  <- solartime(boar_nocattle_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnocattlesummer <- fitact(boar_nocattle_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_nocattle_summer_day<-partial.act(boarnocattlesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nocattle_summer_night<-partial.act(boarnocattlesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_cattle_summer_Ckern <- compareCkern(boarcattlesummer, boarnocattlesummer, reps = 100) # Overlap activity patterns
boar_cattle_summer_Act<- compareAct(list(boarcattlesummer, boarnocattlesummer)) # Difference in activity levels 

# Plot
overlap::densityPlot(cattle_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="C) Wild boar - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_summer$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(boar_cattle_all_summer $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(boar_nocattle_all_summer $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("High cattle abundances","Low cattle abundances", "Cattle", "Shepherd dog"), lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", boar_cattle_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_cattle_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
boar_cattle_summer_direct_Ckern <- compareCkern(boarcattlesummer, cattleallsummer, reps = 100)
boar_nocattle_summer_direct_Ckern <- compareCkern(boarnocattlesummer, cattleallsummer, reps = 100)


# Dog presence
boar_dog_summer <- subset(boar_summer, Dog == 1) # Create data subset
boar_dog_all_summer <- solartime(boar_dog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boardogsummer <- fitact(boar_dog_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_dog_summer_day<-partial.act(boardogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_dog_summer_night<-partial.act(boardogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
boar_nodog_summer  <- subset(boar_summer, Dog == 0) # Create data subset
boar_nodog_all_summer  <- solartime(boar_nodog_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnodogsummer <- fitact(boar_nodog_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_nodog_summer_day<-partial.act(boarnodogsummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nodog_summer_night<-partial.act(boarnodogsummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_dog_summer_Ckern <- compareCkern(boardogsummer, boarnodogsummer, reps = 100) # Overlap activity patterns
boar_dog_summer_Act<- compareAct(list(boardogsummer, boarnodogsummer)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_summer$solar, n.grid=100, ylim=c(0,0.20),main="A) Wild boar - Dog - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(boar_dog_all_summer $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(boar_nodog_all_summer $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("purple3","mediumseagreen", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", boar_dog_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_dog_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
boar_dog_summer_direct_Ckern <- compareCkern(boardogsummer, dogallsummer, reps = 100)
boar_nodog_summer_direct_Ckern <- compareCkern(boarnodogsummer, dogallsummer, reps = 100)



## Winter 

# High cattle abundance
boar_cattle_winter <- subset(boar_winter, Cattle == "high") # Create data subset
boar_cattle_all_winter <- solartime(boar_cattle_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarcattlewinter <- fitact(boar_cattle_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_cattle_winter_day<-partial.act(boarcattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_cattle_winter_night<-partial.act(boarcattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low cattle abundance
boar_nocattle_winter  <- subset(boar_winter, Cattle == "low") # Create data subset
boar_nocattle_all_winter  <- solartime(boar_nocattle_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnocattlewinter <- fitact(boar_nocattle_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_nocattle_winter_day<-partial.act(boarnocattlewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nocattle_winter_night<-partial.act(boarnocattlewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_cattle_winter_Ckern <- compareCkern(boarcattlewinter, boarnocattlewinter, reps = 100) # Overlap activity patterns
boar_cattle_winter_Act <- compareAct(list(boarcattlewinter, boarnocattlewinter)) # Difference in activity levels 

# Plot
overlap::densityPlot(cattle_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="D) Wild boar - Winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(dog_all_winter$solar, add=T, n.grid=100,col="black",lty=2,lwd=2)
overlap::densityPlot(boar_cattle_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(boar_nocattle_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2)
legend("topleft", c("High cattle abundances","Low cattle abundances", "Cattle", "Shepherd dog"), lty=c(1,1,1,2), lwd=2, col=c("purple3","mediumseagreen", "black", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", boar_cattle_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_cattle_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with cattle
boar_cattle_winter_direct_Ckern <- compareCkern(boarcattlewinter, cattleallwinter, reps = 100)
boar_nocattle_winter_direct_Ckern <- compareCkern(boarnocattlewinter, cattleallwinter, reps = 100)


# Dog presence
boar_dog_winter <- subset(boar_winter, Dog == 1) # Create data subset
boar_dog_all_winter <- solartime(boar_dog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boardogwinter <- fitact(boar_dog_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_dog_winter_day<-partial.act(boardogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_dog_winter_night<-partial.act(boardogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Dog absence
boar_nodog_winter  <- subset(boar_winter, Dog == 0) # Create data subset
boar_nodog_all_winter  <- solartime(boar_nodog_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnodogwinter <- fitact(boar_nodog_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_nodog_winter_day<-partial.act(boarnodogwinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nodog_winter_night<-partial.act(boarnodogwinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_dog_winter_Ckern <- compareCkern(boardogwinter, boarnodogwinter, reps = 100) # Overlap activity patterns
boar_dog_winter_Act<- compareAct(list(boardogwinter, boarnodogwinter)) # Difference in activity levels

# Plot
overlap::densityPlot(dog_all_winter$solar, n.grid=100, ylim=c(0,0.20),main="A) Wild boar - Dog - winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(boar_dog_all_winter $solar,  add=T, n.grid=100, lwd=2, col="purple3")
overlap::densityPlot(boar_nodog_all_winter $solar, add=T, n.grid=100,col="mediumseagreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("Dog present","Dog absent", "Herding dog"), lty=c(1,1,1), lwd=2, col=c("purple3","mediumseagreen", "black"), bg="white", cex=1.5)
legend("topright", 
       c(sprintf("Δ = %.2f", boar_dog_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_dog_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with dog
boar_dog_winter_direct_Ckern <- compareCkern(boardogwinter, dogallwinter, reps = 100)
boar_nodog_winter_direct_Ckern <- compareCkern(boarnodogwinter, dogallwinter, reps = 100)


```

### Effects of horses

## Roe deer

```{r}
## Summer 

# High horse abundance
deer_horse_summer <- subset(deer_summer, Horse == "high") # Create data subset
deer_horse_all_summer <- solartime(deer_horse_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deerhorsesummer <- fitact(deer_horse_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_horse_summer_day<-partial.act(deerhorsesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_horse_summer_night<-partial.act(deerhorsesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low horse abundance
deer_nohorse_summer  <- subset(deer_summer, Horse == "low") # Create data subset
deer_nohorse_all_summer  <- solartime(deer_nohorse_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernohorsesummer <- fitact(deer_nohorse_all_summer$solar,sample="data",reps=100) # Fit activity function
deer_nohorse_summer_day<-partial.act(deernohorsesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nohorse_summer_night<-partial.act(deernohorsesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_horse_summer_Ckern <- compareCkern(deerhorsesummer, deernohorsesummer, reps = 100) # Overlap activity patterns
deer_horse_summer_Act<- compareAct(list(deerhorsesummer, deernohorsesummer)) # Difference in activity levels 

# Plot
overlap::densityPlot(horse_all_summer$solar, n.grid=100, ylim=c(0,0.14),main="A) Roe deer - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(deer_horse_all_summer $solar,  add=T, n.grid=100, lwd=2, col="violetred3")
overlap::densityPlot(deer_nohorse_all_summer $solar, add=T, n.grid=100,col="limegreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("High horse abundances","Low horse abundances", "Horse"), lty=c(1,1,1), lwd=2, col=c("violetred3", "limegreen", "black", "black"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", deer_horse_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_horse_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with horse
deer_horse_summer_direct_Ckern <- compareCkern(deerhorsesummer, horseallsummer, reps = 100)
deer_nohorse_summer_direct_Ckern <- compareCkern(deernohorsesummer, horseallsummer, reps = 100)


## Winter 

# High horse abundance
deer_horse_winter <- subset(deer_winter, Horse == "high") # Create data subset
deer_horse_all_winter <- solartime(deer_horse_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deerhorsewinter <- fitact(deer_horse_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_horse_winter_day<-partial.act(deerhorsewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_horse_winter_night<-partial.act(deerhorsewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low horse abundance
deer_nohorse_winter  <- subset(deer_winter, Horse == "low") # Create data subset
deer_nohorse_all_winter  <- solartime(deer_nohorse_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
deernohorsewinter <- fitact(deer_nohorse_all_winter$solar,sample="data",reps=100) # Fit activity function
deer_nohorse_winter_day<-partial.act(deernohorsewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
deer_nohorse_winter_night<-partial.act(deernohorsewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
deer_horse_winter_Ckern <- compareCkern(deerhorsewinter, deernohorsewinter, reps = 100) # Overlap activity patterns
deer_horse_winter_Act <- compareAct(list(deerhorsewinter, deernohorsewinter)) # Difference in activity levels 

# Plot
overlap::densityPlot(horse_all_winter$solar, n.grid=100, ylim=c(0,0.14),main="B) Roe deer - Winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(deer_horse_all_winter $solar,  add=T, n.grid=100, lwd=2, col="violetred3")
overlap::densityPlot(deer_nohorse_all_winter $solar, add=T, n.grid=100,col="limegreen",lwd=2)

abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2)
legend("topleft", c("High horse abundances","Low horse abundances", "Horse"), lty=c(1,1,1), lwd=2, col=c("violetred3", "limegreen", "black", "black"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", deer_horse_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", deer_horse_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with horse
deer_horse_winter_direct_Ckern <- compareCkern(deerhorsewinter, horseallwinter, reps = 100)
deer_nohorse_winter_direct_Ckern <- compareCkern(deernohorsewinter, horseallwinter, reps = 100)

```

## Wild boar

```{r}
## Summer

# High horse abundance
boar_horse_summer <- subset(boar_summer, Horse == "high") # Create data subset
boar_horse_all_summer <- solartime(boar_horse_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarhorsesummer <- fitact(boar_horse_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_horse_summer_day<-partial.act(boarhorsesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_horse_summer_night<-partial.act(boarhorsesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low horse abundance
boar_nohorse_summer  <- subset(boar_summer, Horse == "low") # Create data subset
boar_nohorse_all_summer  <- solartime(boar_nohorse_summer$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnohorsesummer <- fitact(boar_nohorse_all_summer$solar,sample="data",reps=100) # Fit activity function
boar_nohorse_summer_day<-partial.act(boarnohorsesummer, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nohorse_summer_night<-partial.act(boarnohorsesummer, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_horse_summer_Ckern <- compareCkern(boarhorsesummer, boarnohorsesummer, reps = 100) # Overlap activity patterns
boar_horse_summer_Act<- compareAct(list(boarhorsesummer, boarnohorsesummer)) # Difference in activity levels 

# Plot
overlap::densityPlot(horse_all_summer$solar, n.grid=100, ylim=c(0,0.14),main="C) Wild boar - Summer",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(boar_horse_all_summer $solar,  add=T, n.grid=100, lwd=2, col="violetred3")
overlap::densityPlot(boar_nohorse_all_summer $solar, add=T, n.grid=100,col="limegreen",lwd=2)

abline(v=c(6+32/60, 20+36/60), lty=3,lwd=2)
legend("topleft", c("High horse abundances","Low horse abundances", "Horse"), lty=c(1,1,1), lwd=2, col=c("violetred3", "limegreen", "black", "black"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", boar_horse_summer_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_horse_summer_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with horse
boar_horse_summer_direct_Ckern <- compareCkern(boarhorsesummer, horseallsummer, reps = 100)
boar_nohorse_summer_direct_Ckern <- compareCkern(boarnohorsesummer, horseallsummer, reps = 100)


## Winter 

# High horse abundance
boar_horse_winter <- subset(boar_winter, Horse == "high") # Create data subset
boar_horse_all_winter <- solartime(boar_horse_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarhorsewinter <- fitact(boar_horse_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_horse_winter_day<-partial.act(boarhorsewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_horse_winter_night<-partial.act(boarhorsewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Low horse abundance
boar_nohorse_winter  <- subset(boar_winter, Horse == "low") # Create data subset
boar_nohorse_all_winter  <- solartime(boar_nohorse_winter$eventStartTime, 42.01683, -8.184782, 0) # Transform to solar time
boarnohorsewinter <- fitact(boar_nohorse_all_winter$solar,sample="data",reps=100) # Fit activity function
boar_nohorse_winter_day<-partial.act(boarnohorsewinter, c(6+32/60, 20+36/60)*pi/12) # Calculate diurnal activity level
boar_nohorse_winter_night<-partial.act(boarnohorsewinter, c(20+36/60,6+32/60)*pi/12) # Calculate nocturnal activity level

# Estimate overlap and difference in activity patterns
boar_horse_winter_Ckern <- compareCkern(boarhorsewinter, boarnohorsewinter, reps = 100) # Overlap activity patterns
boar_horse_winter_Act <- compareAct(list(boarhorsewinter, boarnohorsewinter)) # Difference in activity levels 

# Plot
overlap::densityPlot(horse_all_winter$solar, n.grid=100, ylim=c(0,0.18),main="D) Wild boar - Winter",col="black",lwd=2, cex.lab=1.5, cex.axis=1.3,cex.main=2)
overlap::densityPlot(boar_horse_all_winter $solar,  add=T, n.grid=100, lwd=2, col="violetred3")
overlap::densityPlot(boar_nohorse_all_winter $solar, add=T, n.grid=100,col="limegreen",lwd=2)

abline(v=c(8+22/60, 18+44/60), lty=3,lwd=2)
legend("topleft", c("High horse abundances","Low horse abundances", "Horse"), lty=c(1,1,1), lwd=2, col=c("violetred3", "limegreen", "black", "black"), bg="white", cex=1.5)
legend("topright",
       c(sprintf("Δ = %.2f", boar_horse_winter_Ckern[["obs"]]),
         sprintf("δ = %.2f", boar_horse_winter_Act[, "Difference"])),
       bg = "white", cex=1.5)

# Estimate overlap with horse
boar_horse_winter_direct_Ckern <- compareCkern(boarhorsewinter, horseallwinter, reps = 100)
boar_nohorse_winter_direct_Ckern <- compareCkern(boarnohorsewinter, horseallwinter, reps = 100)

```

### Generalized Linear Mixed Models (GLMM) - Spatiotemporal effect of livestock


### GLMM analysis (spatiotemporal)

## Load data

```{r}

setwd("H:/Spatiotemporal_analysis/Data")

glmm_data <- read.csv("GLMM_dataset.csv",header=T,sep=";")
ct_loc <- read.csv("CT_locations.csv",header=T,sep=";")

glmm_data <- merge(ct_loc, glmm_data, by="locationID", all.y=T)

glmm_data <- glmm_data %>% separate(locationID, c("loc","year"),sep="_",remove = FALSE)
glmm_data$locationID <- factor(glmm_data$locationID)
glmm_data$loc <- factor(glmm_data$loc)
glmm_data$scientificName<- factor(glmm_data$scientificName)
glmm_data$Season <- factor(glmm_data$Season)

# Adjust core activity times based on summer and winter activity pattern 
glmm_data <- glmm_data %>%
  mutate(CattleActivity = case_when(
    Season == "summer" & Hour >= 7 & Hour <= 22 ~ 1,
    Season == "summer" & (Hour < 7 | Hour > 22) ~ 0,
    Season == "winter" & Hour >= 9 & Hour <= 20 ~ 1,
    Season == "winter" & (Hour < 9 | Hour > 20) ~ 0,
    TRUE ~ NA_real_  # fallback in case something doesn't match
  ))

glmm_data <- glmm_data %>%
  mutate(HorseActivity = case_when(
    Season == "summer" & Hour >= 6 & Hour <= 21 ~ 1,
    Season == "summer" & (Hour < 6 | Hour > 21) ~ 0,
    Season == "winter" & Hour >= 8 & Hour <= 22 ~ 1,
    Season == "winter" & (Hour < 8 | Hour > 22) ~ 0,
    TRUE ~ NA_real_  # fallback in case something doesn't match
  ))

glmm_data$CattleActivity <- factor(glmm_data$CattleActivity)
glmm_data$HorseActivity <- factor(glmm_data$HorseActivity)

# Get season specific encounter rates for cattle and horse from summer and winter dataset
summer_er<- summer %>%
  distinct(locationID, Season, Cattle, Horse, Dog)

winter_er <- winter %>%
  distinct(locationID, Season, Cattle, Horse, Dog)

seasonal_er<- bind_rows(summer_er, winter_er)

# Join with glmm_data to get correct seasonal Cattle and Horse values
glmm_data_updated <- merge(glmm_data,seasonal_er, by =c("locationID","Season"), all.x =T) 
glmm_data_updated$Cattle <- factor(glmm_data_updated$Cattle, levels = c("low", "high"))
glmm_data_updated$Horse <- factor(glmm_data_updated$Horse, levels = c("low", "high"))
glmm_data_updated$Dog <- factor(glmm_data_updated$Dog)

```

## Roe deer - Cattle

```{r}
glmm_deer <- subset(glmm_data_updated,scientificName== "Capreolus capreolus")
hist(cbind(glmm_deer$present,glmm_deer$absent)) 

## Fit model 

glmm_deer_cattle <- glmer(cbind(present,absent) ~ Cattle * CattleActivity * Dog * Season + (1|loc),
                     data = glmm_deer, family = binomial()) #package lme4
summary(glmm_deer_cattle)

## Test for spatial autocorrelation of the residuals using Moran's I

glmm_deer$pr <- residuals(glmm_deer_cattle, type = "pearson")
loc_res <- aggregate(pr ~ loc + lng + lat, data = glmm_deer, mean)
coords <- as.matrix(loc_res[,c("lng","lat")])

nb <- knn2nb(knearneigh(coords, k = 5))
lw <- nb2listw(nb)

moran.test(loc_res$pr, lw)

## Model diagnostics

simulationOutput_deer_cattle <- simulateResiduals(fittedModel = glmm_deer_cattle, plot = F) 
testUniformity(simulationOutput_deer_cattle)
testDispersion(simulationOutput_deer_cattle)
testOutliers(simulationOutput_deer_cattle)
plot(simulationOutput_deer_cattle)

plotResiduals(simulationOutput_deer_cattle, glmm_deer$Cattle)
plotResiduals(simulationOutput_deer_cattle, glmm_deer$Dog)
plotResiduals(simulationOutput_deer_cattle, glmm_deer$Season)

## Effect sizes

# calculate estimated marginal means
em_interaction_deer_cattle <- emmeans(glmm_deer_cattle, ~ Cattle * CattleActivity * Season * Dog, type = "response")

# pairwise contrasts 
pairwise_contrasts_deer_cattle <- contrast(em_interaction_deer_cattle, method = "pairwise", type = "response")
print(pairwise_contrasts_deer_cattle)

# Simple Effects for Cattle Abundance by Cattle Activity, Dog Presence and Season
simple_effects_abundance_deer_cattle <- contrast(em_interaction_deer_cattle, method = "pairwise", by = c("CattleActivity","Season","Dog"))
print(simple_effects_abundance_deer_cattle)

# Simple Effects for Cattle Activity by Cattle Abundance, Dog Presence and Season
simple_effects_activity_deer_cattle <- contrast(em_interaction_deer_cattle, method = "pairwise", by = c("Cattle","Season","Dog"))
print(simple_effects_activity_deer_cattle)

# Simple Effects for Dog Presence by Cattle Abundance, Cattle Activity and Season
simple_effects_dog_deer_cattle <- contrast(em_interaction_deer_cattle, method = "pairwise", by = c("Cattle","CattleActivity","Season"))
print(simple_effects_dog_deer_cattle)

# Convert EMMs to a data frame for plotting
em_interaction_deer_cattle_df <- as.data.frame(em_interaction_deer_cattle)
em_interaction_deer_cattle_df$Cattle = factor(em_interaction_deer_cattle_df$Cattle, levels = c("low", "high"))

# Plot the estimated marginal means 

labels <- c("Summer","Winter")
names(labels) <- c("summer", "winter")

ggplot(em_interaction_deer_cattle_df, 
       aes(x = Cattle, y = prob, color = CattleActivity, 
           linetype = Dog, shape = Dog, 
           group = interaction(CattleActivity, Dog))) +  
  geom_point(position = position_dodge(0.25), size = 4) +
  scale_x_discrete(labels = c("Low", "High")) + 
  scale_color_manual(values = c("mediumseagreen","purple3")) + 
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.5, 
                position = position_dodge(0.25), size = 1) +
  geom_line(position = position_dodge(0.25), size = 1) +
  facet_wrap(~Season, labeller = labeller(Season = labels)) +
  labs(x = "Cattle Abundance",
       y = "Probability of Roe deer Activity",
       color = "Cattle Activity",
       linetype = "Dog",
       shape = "Dog",
       title = "Roe deer - Cattle") +
  theme_bw(base_size = 14) +
  theme(title = element_text(size = 18, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14),
        strip.text.x = element_text(size = 16))

```

## Roe deer - Horse

```{r}
## Fit model

glmm_deer_horse <- glmer(cbind(present,absent) ~ Horse * HorseActivity * Season + (1|loc),
                          data = glmm_deer, family = binomial()) 

summary(glmm_deer_horse)

## Test for spatial autocorrelation of the residuals using Moran's I

glmm_deer$pr <- residuals(glmm_deer_horse, type = "pearson")
loc_res <- aggregate(pr ~ loc + lng + lat, data = glmm_deer, mean)
coords <- as.matrix(loc_res[,c("lng","lat")])

nb <- knn2nb(knearneigh(coords, k = 5))
lw <- nb2listw(nb)

moran.test(loc_res$pr, lw)

## Model diagnostics

simulationOutput_deer_horse <- simulateResiduals(fittedModel = glmm_deer_horse, plot = F) 
testUniformity(simulationOutput_deer_horse)
testDispersion(simulationOutput_deer_horse)
testOutliers(simulationOutput_deer_horse)
plot(simulationOutput_deer_horse)

plotResiduals(simulationOutput_deer_horse, glmm_deer$Horse)
plotResiduals(simulationOutput_deer_horse, glmm_deer$Season)

## Effect sizes

# calculate estimated marginal means
em_interaction_deer_horse <- emmeans(glmm_deer_horse, ~ Horse * HorseActivity * Season, type = "response")

# pairwise contrasts 
pairwise_contrasts_deer_horse <- contrast(em_interaction_deer_horse, method = "pairwise")
print(pairwise_contrasts_deer_horse)

# Simple Effects for Horse by HorseActivity and Season
simple_effects_abundance_deer_horse <- contrast(em_interaction_deer_horse, method = "pairwise", by = c("HorseActivity","Season"))
print(simple_effects_abundance_deer_horse)

# Simple Effects for HorseActivity by Horse and Season
simple_effects_activity_deer_horse <- contrast(em_interaction_deer_horse, method = "pairwise", by = c("Horse","Season"))
print(simple_effects_activity_deer_horse)

# Convert EMMs to a data frame for plotting
em_interaction_deer_horse_df <- as.data.frame(em_interaction_deer_horse)
em_interaction_deer_horse_df$Horse = factor(em_interaction_deer_horse_df$Horse, levels = c("low", "high"))

# Plot the estimated marginal means 

ggplot(em_interaction_deer_horse_df, aes(x = Horse, y = prob, color = HorseActivity, group = HorseActivity)) +  
  geom_point(position = position_dodge(0.2), size = 4) +
  scale_x_discrete(labels = c("Low", "High")) + 
  scale_color_manual(values = c("limegreen","violetred3")) + 
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.5, position = position_dodge(0.2),size=1) + 
  geom_line(position = position_dodge(0.2), size=1) +     
  facet_wrap(~Season, labeller = labeller(Season = labels)) +   
  labs(x = "Horse Abundance", y = "Probability of Roe deer Activity",  
       color = "Horse Activity", 
       title = "Roe deer - Horse") +  
  theme_bw(base_size = 14) +
  theme(title = element_text(size = 18, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"), 
        legend.text = element_text(size = 14), 
        axis.title = element_text(size = 14,face = "bold"),
        axis.text = element_text(size = 14),
        strip.text.x = element_text(size = 16))

```

### Wild boar - Cattle

```{r}
glmm_boar <- subset(glmm_data_updated,scientificName== "Sus scrofa")
hist(cbind(glmm_boar$present,glmm_boar$absent)) 

## Fit model

glmm_boar_cattle <- glmer(cbind(present,absent) ~ Cattle * CattleActivity * Dog * Season + (1|loc),
                     data = glmm_boar, family = binomial()) #package lme4
summary(glmm_boar_cattle)

## Test for spatial autocorrelation of the residuals using Moran's I

glmm_boar$pr <- residuals(glmm_boar_cattle, type = "pearson")
loc_res <- aggregate(pr ~ loc + lng + lat, data = glmm_boar, mean)
coords <- as.matrix(loc_res[,c("lng","lat")])

nb <- knn2nb(knearneigh(coords, k = 5))
lw <- nb2listw(nb)

moran.test(loc_res$pr, lw)

## Model diagnostics

Anova(glmm_boar_cattle)

simulationOutput_boar_cattle <- simulateResiduals(fittedModel = glmm_boar_cattle, plot = F) 
testUniformity(simulationOutput_boar_cattle)
testDispersion(simulationOutput_boar_cattle)
testOutliers(simulationOutput_boar_cattle)
plot(simulationOutput_boar_cattle)

plotResiduals(simulationOutput_boar_cattle, glmm_boar$Cattle)
plotResiduals(simulationOutput_boar_cattle, glmm_boar$Dog)
plotResiduals(simulationOutput_boar_cattle, glmm_boar$Season)

## Effect sizes

# calculate estimated marginal means
em_interaction_boar_cattle <- emmeans(glmm_boar_cattle, ~ Cattle * CattleActivity * Season * Dog, type = "response")

# pairwise contrasts 
pairwise_contrasts_boar_cattle <- contrast(em_interaction_boar_cattle, method = "pairwise")
print(pairwise_contrasts_boar_cattle)

# Simple Effects for Cattle Abundance by Cattle Activity, Dog Presence and Season
simple_effects_abundance_boar_cattle <- contrast(em_interaction_boar_cattle, method = "pairwise", by = c("CattleActivity","Season","Dog"))
print(simple_effects_abundance_boar_cattle)

# Simple Effects for Cattle Activity by Cattle Abundance, Dog Presence and Season
simple_effects_activity_boar_cattle <- contrast(em_interaction_boar_cattle, method = "pairwise", by = c("Cattle","Season","Dog"))
print(simple_effects_activity_boar_cattle)

# Simple Effects for Dog Presence by Cattle Abundance, Cattle Activity and Season
simple_effects_dog_boar_cattle <- contrast(em_interaction_boar_cattle, method = "pairwise", by = c("Cattle","CattleActivity","Season"))
print(simple_effects_dog_boar_cattle)

# Convert EMMs to a data frame for plotting
em_interaction_boar_cattle_df <- as.data.frame(em_interaction_boar_cattle)
em_interaction_boar_cattle_df$Cattle = factor(em_interaction_boar_cattle_df$Cattle, levels = c("low", "high"))

# Plot the estimated marginal means 

labels <- c("Summer","Winter")
names(labels) <- c("summer", "winter")

ggplot(em_interaction_boar_cattle_df, 
       aes(x = Cattle, y = prob, color = CattleActivity, 
           linetype = Dog, shape = Dog, 
           group = interaction(CattleActivity, Dog))) +  
  geom_point(position = position_dodge(0.25), size = 4) +
  scale_x_discrete(labels = c("Low", "High")) + 
  scale_color_manual(values = c("mediumseagreen","purple3")) + 
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.5, 
                position = position_dodge(0.25), size = 1) +
  geom_line(position = position_dodge(0.25), size = 1) +
  facet_wrap(~Season, labeller = labeller(Season = labels)) +
  labs(x = "Cattle Abundance",
       y = "Probability of Wild boar Activity",
       color = "Cattle Activity",
       linetype = "Dog",
       shape = "Dog",
       title = "Wild boar - Cattle") +
  theme_bw(base_size = 14) +
  theme(title = element_text(size = 18, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 14),
        strip.text.x = element_text(size = 16))

```

## Wild boar - Horse

```{r}
## Fit model 

glmm_boar_horse <- glmer(cbind(present,absent) ~ Horse * HorseActivity * Season + (1|loc),
                          data = glmm_boar, family = binomial())
summary(glmm_boar_horse)

## Test for spatial autocorrelation of the residuals using Moran's I

glmm_boar$pr <- residuals(glmm_boar_horse, type = "pearson")
loc_res <- aggregate(pr ~ loc + lng + lat, data = glmm_boar, mean)
coords <- as.matrix(loc_res[,c("lng","lat")])

nb <- knn2nb(knearneigh(coords, k = 5))
lw <- nb2listw(nb)

moran.test(loc_res$pr, lw)

## Model diagnostics

simulationOutput_boar_horse <- simulateResiduals(fittedModel = glmm_boar_horse, plot = F) 
testUniformity(simulationOutput_boar_horse)
testDispersion(simulationOutput_boar_horse)
testOutliers(simulationOutput_boar_horse)
plot(simulationOutput_boar_horse)

plotResiduals(simulationOutput_boar_horse, glmm_boar$horse)
plotResiduals(simulationOutput_boar_horse, glmm_boar$Season)

## Effect sizes

# calculate estimated marginal means
em_interaction_boar_horse <- emmeans(glmm_boar_horse, ~ Horse * HorseActivity * Season, type = "response")

# pairwise contrasts 
pairwise_contrasts_boar_horse <- contrast(em_interaction_boar_horse, method = "pairwise")
print(pairwise_contrasts_boar_horse)

# Simple Effects for Horse by HorseActivity and Season
simple_effects_abundance_boar_horse <- contrast(em_interaction_boar_horse, method = "pairwise", by = c("HorseActivity","Season"))
print(simple_effects_abundance_boar_horse)

# Simple Effects for HorseActivity by Horse and Season
simple_effects_activity_boar_horse <- contrast(em_interaction_boar_horse, method = "pairwise", by = c("Horse","Season"))
print(simple_effects_activity_boar_horse)

# Convert EMMs to a data frame for plotting
em_interaction_boar_horse_df <- as.data.frame(em_interaction_boar_horse)
em_interaction_boar_horse_df$Horse = factor(em_interaction_boar_horse_df$Horse, levels = c("low", "high"))


# Plot the estimated marginal means 
ggplot(em_interaction_boar_horse_df, aes(x = Horse, y = prob, color = HorseActivity, group = HorseActivity)) +  
  geom_point(position = position_dodge(0.2), size = 4) +
  scale_x_discrete(labels = c("Low", "High")) + 
  scale_color_manual(values = c("limegreen","violetred3")) + 
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.5, position = position_dodge(0.2),size=1) + 
  geom_line(position = position_dodge(0.2), size=1) +     
  facet_wrap(~Season, labeller = labeller(Season = labels)) +   
  labs(x = "Horse Abundance", y = "Probability of Wild boar Activity",  
       color = "Horse Activity", 
       title = "Wild boar - Horse") +  
  theme_bw(base_size = 14) +
  theme(title = element_text(size = 18, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"), 
        legend.text = element_text(size = 14), 
        axis.title = element_text(size = 14,face = "bold"),
        axis.text = element_text(size = 14),
        strip.text.x = element_text(size = 16))

```